Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-04-12 emsoft-2019/abstract.tex
adding missing numbers to introduction

We propose an optimization for space-efficient implementations of explicit model-predictive controllers (MPC)
for robust control of linear time-invariant (LTI) systems on embedded platforms. 
We obtain an explicit-form robust model-predictive controller as a solution to a multi-parametric 
linear programming problem.
The structure of the controller is a polyhedral decomposition of the control domain,
with an affine map for each domain.
While explicit MPC is suited for embedded devices with low computational power, the memory requirements
for such controllers can be high.
We provide an optimization algorithm for a mixed-precision implementation of the controller,
where the deviation of the implemented controller from the original one is within the robustness
margin of the robust control problem.
The core of the mixed-precision optimization is an iterative static analysis that co-designs
a robust controller and a low-bitwidth approximation 
that is statically guaranteed to always be within the robustness margin to the original controller.
We have implemented our algorithm and show, on a set of benchmarks, that our optimization
can reduce space requirements by up to 20\% (and about 10\% on average), compared to a uniform precision
implementation of the original controller.

>>>>

We propose an optimization for space-efficient implementations of explicit model-predictive controllers (MPC)
for robust control of linear time-invariant (LTI) systems on embedded platforms. 
We obtain an explicit-form robust model-predictive controller as a solution to a multi-parametric 
linear programming problem.
The structure of the controller is a polyhedral decomposition of the control domain,
with an affine map for each domain.
While explicit MPC is suited for embedded devices with low computational power, the memory requirements
for such controllers can be high.
We provide an optimization algorithm for a mixed-precision implementation of the controller,
where the deviation of the implemented controller from the original one is within the robustness
margin of the robust control problem.
The core of the mixed-precision optimization is an iterative static analysis that co-designs
a robust controller and a low-bitwidth approximation 
that is statically guaranteed to always be within the robustness margin of the original controller.
We have implemented our algorithm and show on a set of benchmarks that our optimization
can reduce space requirements by up to 20.9\% and on average by 12.6\% 
compared to a minimal uniform precision implementation of the original controller.

### 2019-04-12 emsoft-2019/intro.tex
adding missing numbers to introduction

% !TEX root = main.tex
\section{Introduction}
\eva{General comment: it feels a bit long until the intro discusses the actual problem,
but it reads nicely and motivates the problem well, so it's probably fine}
\eva{General comment: should we add some more references to related work (and not only to MPC)?}
Model predictive control (MPC) is a technique to design control actions by solving finite-horizon open-loop
optimal control problems at each sampling instant.
The result of each optimization gives a sequence of optimal control actions, only the first of which is applied
to the process.
The same procedure is applied in the next time instant with a shifted time horizon and a new initial state, 
after receiving the updated values of the process state.
The optimization problem in MPC uses a dynamic model of the process, encodes all input and output (state) constraints, and 
optimizes a performance index. 
MPC has shown to be successful in a wide variety of industrial applications, due to its 
ability to systematically handle processes with many state and input variables as well as constraints on them.

@@

However, there are still bottlenecks in implementing explicit MPC on resource-constrained embedded micro-controllers.
First, implementations of explicit MPCs can suffer from large memory usage, because the solution of the optimization
problem can involve many (often hundreds) of regions.
Since the memory consumption grows linearly with the number of regions, 
this can be a limited factor in using explicit MPC on resource-constrained micro-controllers.
Second, many low-end micro-controllers only support fixed-point arithmetic.
Errors in the controller implementation are inversely proportional to the number of bits used for representing each variable. 
An implementation of explicit MPC has to be \emph{robust} to 
implementation errors to be able to enforce hard constraints on the states at run time.

In this paper, we consider the problem of implementing explicit MPC on low-end microcontrollers with fixed-point arithmetic
in a \emph{memory-efficient} and \emph{robust} way.
We propose an automatic controller and mixed-precision implementation co-design technique that computes a minimal set \todo{What is meant by minimal?} of mixed
precision assignments for all variables while ensuring the resulting implementation error remains within the robustness margin.

Our proposed method iteratively solves a robust version of the MPC which explicitly considers fixed-point implementation errors
in the model; see~\autoref{fig:overview}.
Initially, we estimate a bound $\Delta = \Delta_0$ on the implementation error and solve a
min-max quadratic program for explicit robust MPC \cite{delaPea:2005}, where the system
model has an explicit disturbance bounded by $\Delta$.
The solution is represented as a piecewise affine map.
We use an automated mixed-precision tuning tool to find an approximate implementation of the piecewise affine
map. 
We statically compute the error of the implementation, taking into account errors
in evaluating the affine maps as well as choosing the wrong affine map due to quantization.
If the error is at most $\Delta$, we are done: the implementation satisfies the robustness margin in the model.
If not, we increase the error bound $\Delta$ and run the loop again to find a new robust controller and find its best mixed-precision
implementation.

We have implemented our algorithm on top of Matlab's multi-parametric toolbox 
for robust explicit MPC~\cite{matlabMPT}
and the Daisy tool for multi-precision tuning and fixed-point error analysis \cite{Daisy}.
We have applied our technique on a number of standard benchmark examples.
Our implementation finds that our mixed-precision implementation can save up to 20\% memory (on average, \RM{XXX}\%) in controller
implementations over a uniform-precision implementation (and a saving of \RM{XXX}\% over a uniform 32-bit implementation), while
maintaining the correctness of the controller.
In absolute terms, the saving corresponds to \RM{YYY} KB over a uniform precision implementation on our largest benchmark.
In each case, the analysis takes only a few minutes of computation.
We also demonstrate the scalability of our mixed-precision tuning and error bounds analysis:
we show that on explicit MPC controllers with thousands of regions, our tool finishes in a few minutes\todo{Hours?}.

We summarize our contributions.
\begin{itemize}
\item We provide an end-to-end tuning algorithm for robust explicit MPC;
\item We design a static and scalable error analysis algorithm for piecewise affine functions;
\item Through a set of standard control benchmarks, we demonstrate that our tuning algorithm
can achieve up to 20\% savings in memory while maintaining robustness.
\end{itemize}

>>>>

% !TEX root = main.tex
\section{Introduction}
\eva{General comment: should we add some more references to related work (and not only to MPC)?}
Model predictive control (MPC) is a technique to design control actions by solving finite-horizon open-loop
optimal control problems at each sampling instant.
The result of each optimization gives a sequence of optimal control actions, only the first of which is applied
to the process.
The same procedure is applied in the next time instant with a shifted time horizon and a new initial state, 
after receiving the updated values of the process state.
The optimization problem in MPC uses a dynamic model of the process, encodes all input and output (state) constraints, and 
optimizes a performance index. 
MPC has shown to be successful in a wide variety of industrial applications, due to its 
ability to systematically handle processes with many state and input variables as well as constraints on them. 
\todo{any citations?}

@@

However, there are still bottlenecks in implementing explicit MPC on resource-constrained embedded micro-controllers.
First, implementations of explicit MPCs can suffer from large memory usage, because the solution of the optimization
problem can involve many (often hundreds) of regions. \todo{citation for such example, e.g. the double integrator}
Since the memory consumption grows linearly with the number of regions, 
this can be a limiting factor in using explicit MPC on resource-constrained micro-controllers.
Second, many low-end micro-controllers only support fixed-point arithmetic.
Errors in the controller implementation are inversely proportional to the number of bits used for representing each variable~\cite{Anta10}. 
An implementation of explicit MPC has to be \emph{robust} to 
implementation errors to be able to enforce hard constraints on the states at run time~\cite{imperialrmpc}.

In this paper, we consider the problem of implementing explicit MPC on low-end microcontrollers with fixed-point arithmetic
in a \emph{memory-efficient} and \emph{robust} way.
We propose an automatic controller and mixed-precision implementation co-design
technique that computes mixed-precision assignments which minimize bitwidths for all
variables while ensuring the resulting implementation error remains within the
robustness margin.

Our proposed method iteratively solves a robust version of MPC which explicitly considers finite-precision implementation errors
in the model; see~\autoref{fig:overview}.
Initially, we estimate a bound $\Delta = \Delta_0$ on the implementation error and solve a
min-max quadratic program for explicit robust MPC \cite{delaPea:2005}, where the system
model has an explicit disturbance bounded by $\Delta$.
The solution is represented as a piecewise affine map.
We use an automated mixed-precision tuning tool to find an approximate implementation of the piecewise affine
map. 
We statically compute the error of the implementation, taking into account errors
in evaluating the affine maps as well as choosing the wrong affine map due to quantization.
If the error is at most $\Delta$, we are done: the implementation satisfies the robustness margin in the model.
If not, we increase the error bound $\Delta$ and run the loop again to find a new robust controller and find its best mixed-precision
implementation.

To the best of our knowledge, our algorithm is the first to automatically and soundly
synthesize a suitable fixed-point precision for explicit robust MPC controllers,
and the first to consider mixed-precision in this context.

We have implemented our algorithm on top of Matlab's multi-parametric toolbox 
for robust explicit MPC~\cite{matlabMPT}
and the Daisy tool for multi-precision tuning and fixed-point error analysis~\cite{Daisy}.
We have applied our technique on a number of standard benchmark examples.

Our algorithm finds mixed-precision implementations which save up to 20.9\%
memory (on average, 12.6\%) in controller implementations over a minimal
uniform-precision implementation (and an average saving of 46.8\% over a uniform
32-bit implementation), while maintaining the correctness of the controller.

For smaller examples, the analysis takes only a few minutes of computation.
We also demonstrate the scalability of our mixed-precision tuning and error bounds analysis:
we show that on explicit MPC controllers with thousands of regions, our tool finishes in a few hours.
In absolute terms, the memory saving corresponds to 21KB over a uniform
precision implementation on our largest benchmark.

In summary, our contributions are:
\begin{itemize}
\item We design a static and scalable error analysis algorithm for piecewise affine functions;
\item We provide an end-to-end tuning algorithm for robust explicit MPC;
\item Through a set of standard control benchmarks, we demonstrate that our tuning algorithm
can achieve up to 20\% savings in memory while maintaining robustness.
\end{itemize}
